---
title: "Ch02"
author: "Vincent Alegrete"
date: "2024-12-22"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1 (Top Base Stealers in the Hall of Fame)

The following table gives the number of stolen bases (SB), the number of times caught stealing (CS), and the number of games played (G) for nine players currently inducted in the Hall of Fame.

(a) In R, place the stolen base, caught stealing, and game counts in the vectors <tt>SB</tt>, <tt>CS</tt>, and <tt>G</tt>.

<i>Solution</i>

```{r}
SB <- c(1406, 938, 897, 741, 738, 689, 506, 504, 474)
CS <- c(335, 307, 212, 195, 109, 162, 136, 131, 114)
G <- c(3081, 2616, 3034, 2826, 2476, 2649, 2599, 2683, 2379)
```

(b) For all players, compute the number of stolen base attempts <tt>SB + CS</tt> and store in the vector <tt>SB.Attempt</tt>.

<i>Solution</i>

```{r}
SB.Attempt <- SB + CS
```

(c) For all players, compute the success rate <tt>Success.Rate = SB/SB.Attempt</tt>.

<i>Solution</i>

```{r}
Success.Rate <- SB / SB.Attempt
```

(d) Compute the number of stolen bases per game <tt>SB.Game = SB/Game</tt>.

<i>Solution</i>

```{r}
SB.Game <- SB / G
```

(e) Construct a scatterplot of the stolen bases per game against the success rates. Are there particular players with unusually high or low stolen base success rates? Which player had the greatest number of stolen bases per game?

<i>Solution</i> The plot does show some unusually high and low stolen base success rates:

```{r}
plot(SB.Game ~ Success.Rate)
```

To get the requested players, we cross-reference a vector of player names:

```{r}
Names <- c("Rickey Henderson", "Lou Brock", "Ty Cobb", "Eddie Collins",
           "Max Carey", "Joe Morgan", "Luis Aparicio", "Paul Molitor",
           "Roberto Alomar")

Names[which.min(Success.Rate)]
Names[which.max(Success.Rate)]
Names[which.max(SB.Game)]
```

### Problem 2 (Character, Factor, and Logical Variables in R)

Suppose one records the outcomes of a batter in ten plate appearances: Single, Out, Out, Single, Out, Double, Out, Walk, Out, Single.

(a) Use the <tt>c</tt> function to collect these outcomes in a character vector <tt>outcomes</tt>.

```{r}
outcomes <- c("Single", "Out", "Out", "Single", "Out",
              "Double", "Out", "Walk", "Out", "Single")
```

(b) Use the table function to construct a frequency table of outcomes.

```{r}
table(outcomes)
```

(c) In tabulating these last results, suppose one prefers the results to be ordered from least-successful to most-successful. Convert the character vector outcomes to a factor variable. Use the table function to tabulate the values in <tt>f.outcomes</tt>. How does the output differ from what you saw in part (b)?

<i>Solution</i> The table is now ordered according to our factor levels:

```{r}
f.outcomes <- factor(outcomes,
                     levels=c("Out", "Walk", "Single", "Double"))
table(f.outcomes)
```

(d) Suppose you want to focus only on the walks in the plate appearances. Describe what is done in each of the following statements:

<tt>outcomes == "Walk"</tt>

<tt>sum(outcomes == "Walk")</tt>

<i>Solution</i> The first line returns a boolean vector containing the elementwise results of this test:

```{r}
outcomes == "Walk"
```

Summing the vector gives the number of walks (<tt>TRUE</tt> is treated as a 1):
```{r}
sum(outcomes == "Walk")
```


# Problem 3 (Pitchers in the 350-Wins Club)